Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-05-29 manuscript.Rmd
Fix in-line image issue

Social coordination has long been used both physically and virtually
to start and carry out collective actions. These collective events
bring individuals together and create emergent patterns of communication
and action. From a dynamical
systems perspective, these different means of social coordination during
times of turmoil can be considered two complementary components of a
single system, mutually constrained and constantly interacting and
influencing one another [@mahner1997foundations]. 
Increasingly, social media platforms are being utilized during
social mobilization, but the relationship between virtual and real-world
mobilization is still largely unknown. Investigation of the
coupling of real-world mobilization and social cohesion could provide
new insights into the power of social media during mobilization and 
it can be used to respond to or create social change. The present work 
aims to understand how physical and virtual movements coexist and interact 
from a dynamical systems perspective.

@@

## Windowed Cross-Recurrence Quantification Analysis
```{r primary-wcrqa, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}

>>>>

Social coordination has long been used both physically and virtually
to start and carry out collective actions. These collective events
bring individuals together and create emergent patterns of communication
and action. Online communication has historically been used to document
and communicate archival information about events, but it has been
increasingly used to bring about social change. From a dynamical
systems perspective, these different means of social coordination during
times of turmoil can be considered two complementary components of a
single system, mutually constrained and constantly interacting and
influencing one another [@mahner1997foundations]. Investigation of the
coupling of real-world mobilization and social cohesion could provide
new insight into the power of social media during mobilization and the
ways in which individuals utilize it to bring about, or respond to, social
change. Increasingly, social media platforms are being utilized during
social mobilization, but the relationship between virtual and real-world
mobilization is still largely unknown. The present work aims to understand
how physical and virtual movements coexist and interact from a dynamical
systems perspective.

@@

## Windowed Cross-Recurrence Quantification Analysis
```{r primary-wcrqa, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide', include = FALSE}

### 2020-05-29 manuscript.tex
Fix in-line image issue

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring},
            pdfkeywords={Social cohesion, recurrence quantification analysis, complex systems, Twitter, nonlinear dynamics, Arab Spring},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

Correspondence concerning this article should be addressed to Megan Chiovaro, 406 Babbidge Road, Unit 1020, Storrs, CT 06269. E-mail: megan.chiovaro@uconn.edu}
\keywords{Social cohesion, recurrence quantification analysis, complex systems, Twitter, nonlinear dynamics, Arab Spring\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\abstract{
Social mobilization requires effective collective coordination. In recent years, individuals involved in social movements have taken to social media platforms such as Twitter to reach broader audiences at rapid speeds. Effective virtual mobilization similarly requires social cohesion and coordination, yet it is still unclear if virtual mobilization reflects real-world movements. Here, we explore the dynamics of real-world events and social cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear methods cross-recurrence quantification analysis and windowed cross-recurrence quantification analysis, we investigate if counts of all daily events, daily positive events, and daily negative events are coupled with social cohesion found in Syrian tweets. Results indicate that virtual social cohesion is coupled with the counts of all and negative events each day, but not with counts of positive events. This suggests that virtual mobilization does reflect real-world movements, but only for salient, frequent and negative events. We conclude with a discussion of implications and applications of nonlinear methods in political science research.
}

@@

Traditional linear analyses have led to many notable discoveries in the domains of psychology and political science, but continuous data pose problems that are not easily overcome with these methods. Events have historically been conceptualized as linear chains of smaller, sequential events with one occurrence leading to the next. This notion has been increasingly replaced by ideas from dynamical systems theory. Rather than a chain of events, continuous data can be thought of as nested occurrences, where mutually constrained parts (e.g., citizens, governments, real-world events, social media platforms) are constantly interacting, leading to more global behaviors that are larger than the sum of the parts (Turvey, 2018). When analyzed as dynamical systems, many phenomena exhibit new nonlinear patterns that were formerly overlooked. Continuous data subject to linear analyses becomes flattened, and the variability that is fundamental to these systems and their evolution over time is ignored. Nonlinear methods provide means for capturing the rich temporal dynamics and variability that unfold in continuous data.
The primary methods used in the current work is cross-recurrence quantification analysis (CRQA; Zbilut, Giuliani, \& Webber Jr, 1998) and it's off-shoot, windowed cross-recurrence quantification analysis (WCRQA; Webber Jr \& Zbilut, 2005). CRQA has successfully been used to gain a better understanding of many complex human phenomena, such as human social interactions (for a review, see Fusaroli, Konvalinka, \& Wallot, 2014) and interpersonal synchrony (Paxton \& Dale, 2017). We argue that the use of nonlinear methods in the investigation of political science phenomena will lead to new insights and development of more robust theories.

@@

The corpus of tweets used included 697428 Syrian tweets occurring between March 30, 2012 and June 15, 2012. Using the Twitter streaming API, servers at the University of Texas at Austin collected the tweets. The servers were run on a stable (LTS) version of Ubuntu Linus, 10.04, and were connected directly to the high-bandwidth UT network maintained by IT services.
Tweets collected contained one or more of the following keywords: \enquote{syria,} \enquote{syrian,} \enquote{damascus,} \enquote{homs,} \enquote{al-assad,} and \enquote{sunni.} These keywords were chosen based on the most distinctive terms used in news articles covering the Arab Spring in Syria at the time. A node.js script was used to reinitialize the connection to Twitter's servers whenever the script timed out due to a period of silence. The script requested data from the API endpoint using basic HTTP authentication (OAuth was optional, at the time) at \url{https://stream.twitter.com/1.1/statuses/filter.json}, and stored the data directly to the local hard disk. The tweets were then post-processed with a Python script to convert the JSON-formatted tweets to tab-separated files and were then compressed to maximize storage efficiency.
All scripts used for scrapping tweets and compressing data can be found at \url{https://github.com/chbrown/twilight}.

@@

The social cohesion or linguistic similarity of daily tweets was calculated by quantifying the frequency of shared content words with respect to those words' frequencies in the entire tweet corpus. If a linguistic term has a high frequency in one tweet but has a low frequency in the whole tweet corpus, then it is more likely a content word and is this given more weight in calculating social cohesion.
The tweets were first sorted into ascending order by timestamp. They were then grouped into successive sets of five tweets. Inverse frequency weighting for the overlap words and all words was conducted for each pair of two tweets in the group of five, then divided to get the social cohesion score for that pair (see Eq. 1-3). This resulted in ten social cohesion values for each group of five tweets, which were then averaged. Finally, the mean of all average cohesion values for all groups was taken to get a social cohesion score for each day.

@@

Using the Integrated Crisis Early Warning System (ICEWS) developed at the Defense Advanced Research Project Agency, the daily events in Syria were scored with an intensity rating ranging from -10 to 10. Negative events, such as riots and attacks, are coded with low negative numbers while positive, peaceful events are coded with high positive numbers. This well-validated coding scheme allows for numerical quantification of different event types.
For both the target only and the target and source data, time series of the count of all events were generated by summing the total number of events on each day, irrespective of ICEWS score. Similarly, time series of count of positive events were generated by summing the number of ICEWS events for each day that had an intensity rating greater than zero. Lastly, time series of count of negative events were generated by summing the number of ICEWS events for each day that had an intensity rating less than zero.

@@

We first used cross-recurrence quantification analysis (CRQA; Zbilut et al., 1998) to investigate the temporal patterns of social cohesion and real-world events in Syria during Arab Spring. As an extension of RQA, which captures the patterns of revisited states of a single time series (Webber Jr \& Zbilut, 1994; Zbilut \& Webber Jr, 1992), CRQA allows for two time series to be projected onto one another to investigate shared states and trajectories. Nonlinear analyses such as CRQA can reveal dynamics that are not captured by traditional linear methods. In order to perform CRQA, the time series need to be converted to the same scale, while still preserving a sampling rate that respects the native timescale of the system (cf.~Nyquist frequency in signal processing; Grenander, 1959). We converted each time series (e.g., Figure \ref{fig:raw-ts}) into deciles (e.g., Figure \ref{fig:deciled-ts}) to analyze the dynamics of different levels of social cohesion and counts of all, positive, and negative events on each day. In CRQA, a recurrent point is when two signals share the same state. These shared states can be at the same time or at later points in the time series.

@@

The percentage of recurrent points to total possible points is a metric called recurrence rate (RR). This metrics gives a general description of how frequently the time series are sharing the same states. In the current study, a recurrent point would indicate that the level of social cohesion and frequency of events were at the same intensity.
Recurrent points that occur in successive moments form line structures. These are conceptualized as shared trajectories, where the two time series are moving together across time in a shared state. The percentage of recurrent points that make up diagonal line structures (i.e., two or more consecutive points) is a metric called percent determinism (DET). For the current project, a high DET would indicate that not only are the levels of social cohesion and frequency of events occurring at similar intensities, but they are doing so across consecutive days.
The average length of the line structures is called mean line (L). This can be conceptualized as the average amount of time that the two time series move together before one or both shift states. The longest line length is called max line (maxL). maxL indicates the longest instance of being \enquote{stuck} in a particular state together. The total number of lines is called NRLINE. The more lines there are, the more coordinated the system may be. The complexity or variability of line lengths is captured by a metric called entropy (ENTR) and entropy while accounting for the total number of lines in the plot is normalized entropy (rENTR). These entropy metrics help indicate the stability and predictability of the system. When one of the signals gets \enquote{stuck} in a state and the other moves on to other states, this results in vertical line structures. The proportion of points forming vertical line structures is a metric called laminarity (LAM) and the average length of the vertical trajectories is called trapping time (TT). For more detail on CRQA metrics, see Coco and Dale (2018).

@@

The time period of differing dynamics, from observation 38 through 43, appears to be captured by WCRQA of social cohesion and count of negative events for both RR and DET (Figure \ref{fig:plot-wcrqa-targ-all}). The amount of recurrence and determinism for these periods almost consistently lie above chance, suggesting that the intensity of negative events coincide with the degree of social cohesion found on Twitter, and that these states carry across multiple days in a way that is different from that of chance.
The above-chance levels of determinism and recurrence rate are demonstrating the coupling of social cohesion and counts of all and negative events during times of strife. We can see this string of highly violent events reflected in the mode ICEWS event for each day (Figure \ref{fig:plot-mode-event}), where from May 1st, 2020 to May 15th, 2020, the most frequently occurring event was almost consistently -10, the most violent event in the coding system. These dates directly correlate with the above chance RR and DET metrics for the WCRQA of social cohesion and count of negative events. Furthermore, for WCRQA of social cohesion and positive event count during this time period, DET was entirely within chance (Figure \ref{fig:plot-wcrqa-targ-all}). This supports our hypothesis, as the positive events are seemingly not coupled with the social cohesion on Twitter during civil uprisings during the Arab Spring.

@@

\begin{verbatim}
## Warning in readLines(x, encoding = "UTF-8"): incomplete final line found on
## 'appendix.Rmd'
\end{verbatim}

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\hypertarget{crqa-for-phoenix-data-in-sextiles}{%
\section{CRQA for Phoenix data in
sextiles}\label{crqa-for-phoenix-data-in-sextiles}}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for target only Phoenix data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 31.065 & 0.241 & & 38.955 & 0.000 & *** & 31.854 & 0.139
&\tabularnewline
NRLINE & 144.000 & 0.229 & & 174.000 & 0.000 & *** & 148.000 & 0.105
&\tabularnewline
maxL & 4.000 & 0.421 & & 5.000 & 0.087 & . & 4.000 & 0.452
&\tabularnewline
L & 2.188 & 0.544 & & 2.270 & 0.039 & * & 2.182 & 0.580 &\tabularnewline
ENTR & 0.513 & 0.526 & & 0.650 & 0.033 & * & 0.504 & 0.550
&\tabularnewline
rENTR & 0.467 & 0.290 & & 0.469 & 0.276 & & 0.459 & 0.310
&\tabularnewline
LAM & 32.051 & 0.284 & & 46.154 & 0.008 & ** & 30.769 & 0.314
&\tabularnewline
TT & 2.273 & 0.196 & & 2.400 & 0.072 & . & 2.400 & 0.063 &
.\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for source and target Phoenix
data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 31.558 & 0.178 & & 37.377 & 0.000 & *** & 25.740 & 0.988
&\tabularnewline
NRLINE & 145.000 & 0.189 & & 169.000 & 0.000 & *** & 119.000 & 0.989
&\tabularnewline
maxL & 4.000 & 0.467 & & 5.000 & 0.084 & . & 4.000 & 0.454
&\tabularnewline
L & 2.207 & 0.368 & & 2.243 & 0.123 & & 2.193 & 0.509 &\tabularnewline
ENTR & 0.547 & 0.343 & & 0.610 & 0.104 & & 0.521 & 0.492
&\tabularnewline
rENTR & 0.498 & 0.135 & & 0.440 & 0.432 & & 0.474 & 0.244
&\tabularnewline
LAM & 38.462 & 0.073 & . & 48.718 & 0.000 & *** & 21.795 & 0.790
&\tabularnewline
TT & 2.500 & 0.017 & * & 2.235 & 0.280 & & 2.429 & 0.048 &
*\tabularnewline
\bottomrule
\end{longtable}

CRQ analyses for sextiled Phoenix event data found increased levels of
coupling between positive events and social cohesion, irrespective of
target or source and target filtering. Negative events were not coupled
with social cohesion irrespective of filtering.

\hypertarget{crqa-for-icews-data-in-sextiles}{%
\section{CRQA for ICEWS data in
sextiles}\label{crqa-for-icews-data-in-sextiles}}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for target only ICEWS data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 39.250 & 0.000 & *** & 37.870 & 0.000 & *** & 38.659 & 0.000 &
***\tabularnewline
NRLINE & 177.000 & 0.000 & *** & 169.000 & 0.000 & *** & 170.000 & 0.000
& ***\tabularnewline
maxL & 5.000 & 0.093 & . & 5.000 & 0.081 & . & 5.000 & 0.087 &
.\tabularnewline
L & 2.249 & 0.089 & . & 2.272 & 0.033 & * & 2.306 & 0.009 &
**\tabularnewline
ENTR & 0.619 & 0.074 & . & 0.653 & 0.030 & * & 0.705 & 0.008 &
**\tabularnewline
rENTR & 0.447 & 0.368 & & 0.471 & 0.261 & & 0.509 & 0.118
&\tabularnewline
LAM & 43.590 & 0.013 & * & 39.744 & 0.047 & * & 42.308 & 0.030 &
*\tabularnewline
TT & 2.267 & 0.208 & & 2.385 & 0.065 & . & 2.357 & 0.092 &
.\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for source and target ICEWS data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 34.122 & 0.014 & * & 37.771 & 0.000 & *** & 39.448 & 0.000 &
***\tabularnewline
NRLINE & 152.000 & 0.064 & . & 168.000 & 0.000 & *** & 173.000 & 0.000 &
***\tabularnewline
maxL & 5.000 & 0.092 & . & 5.000 & 0.093 & . & 7.000 & 0.002 &
**\tabularnewline
L & 2.276 & 0.030 & * & 2.280 & 0.024 & * & 2.312 & 0.003 &
**\tabularnewline
ENTR & 0.663 & 0.022 & * & 0.661 & 0.022 & * & 0.698 & 0.005 &
**\tabularnewline
rENTR & 0.478 & 0.211 & & 0.477 & 0.242 & & 0.434 & 0.475
&\tabularnewline
LAM & 34.615 & 0.178 & & 39.744 & 0.047 & * & 48.718 & 0.002 &
**\tabularnewline
TT & 2.250 & 0.233 & & 2.385 & 0.067 & . & 2.235 & 0.295
&\tabularnewline
\bottomrule
\end{longtable}

CRQ analyses for sextiled ICEWS event data found coupling between both
positive and negative events with social cohesion, irrespective of
target or source and target filtering. Negative events were more
strongly coupled with social cohesion than positive events for the
target only analyses. The absence of coupling between negative events
and social cohesion in the Phoenix CRQ analyses speaks to the
unreliability of event data, and that nonlinear analyses in general are
only as valuable as the quality of the data being used.

\hypertarget{windowed-crqa-for-phoenix-data-in-sextiles}{%
\section{Windowed CRQA for Phoenix data in
sextiles}\label{windowed-crqa-for-phoenix-data-in-sextiles}}

\begin{figure}
\includegraphics[width=0.9\linewidth,angle=90]{./results/secondary/phoenix/windowed-crqa/target-windowed_all} \caption{Windowed CRQA for social cohesion and event counts from target filtered data.}\label{fig:plot-secondary-phoenix-wcrqa-targ-all}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth,angle=90]{./results/secondary/phoenix/windowed-crqa/source_target-windowed_all} \caption{Windowed CRQA for social cohesion and event counts from target filtered data.}\label{fig:plot-secondary-phoenix-wcrqa-source-targ-all}
\end{figure}

\hypertarget{windowed-crqa-for-icews-data-in-sextiles}{%
\section{Windowed CRQA for ICEWS data in
sextiles}\label{windowed-crqa-for-icews-data-in-sextiles}}

\begin{figure}
\includegraphics[width=0.9\linewidth,angle=90]{./results/secondary/icews/windowed-crqa/target-windowed_all} \caption{Windowed CRQA for social cohesion and event counts from target filtered data.}\label{fig:plot-secondary-icews-wcrqa-targ-all}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth,angle=90]{./results/secondary/icews/windowed-crqa/source_target-windowed_all} \caption{Windowed CRQA for social cohesion and event counts from target filtered data.}\label{fig:plot-secondary-icews-wcrqa-source-targ-all}
\end{figure}
\end{appendix}

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring},
            pdfkeywords={social cohesion, recurrence quantification analysis, complex systems, Twitter, Arab Spring},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

Correspondence concerning this article should be addressed to Megan Chiovaro, 406 Babbidge Road, Unit 1020, Storrs, CT 06269. E-mail: megan.chiovaro@uconn.edu}
\keywords{social cohesion, recurrence quantification analysis, complex systems, Twitter, Arab Spring\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\abstract{
Social mobilization requires effective collective coordination. In recent
years, individuals involved in social movements have taken to social media
platforms to reach broader audiences at rapid speeds. Effective virtual
mobilization similarly requires social cohesion and coordination, yet the
degree to which virtual mobilization reflects or drives real-world movements
is unclear. Here, we explore the dynamics of real-world events and social
cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear
methods cross-recurrence quantification analysis and windowed cross-recurrence
quantification analysis, we investigate if counts of all daily events, daily
positive events, and daily negative events are coupled with social cohesion
found in Syrian tweets. Results indicate that virtual social cohesion is
coupled with the counts of all and negative events each day but not with
counts of positive events. This suggests that virtual mobilization does reflect
real-world movements but only for salient, frequent and negative events. We
conclude with a discussion of implications and applications of nonlinear
methods in political science research.
}

@@

Traditional linear analyses have led to many notable discoveries in the
domains of psychology and political science, but continuous data pose
problems that are not easily overcome with these methods. Events have
historically been conceptualized as linear chains of smaller, sequential
events with one occurrence leading to the next. This notion has been
increasingly replaced by ideas from dynamical systems theory. Rather than
a chain of events, continuous data can be thought of as nested occurrences,
where mutually constrained parts (e.g., citizens, governments, real-world
events, social media platforms) are constantly interacting, leading to
more global behaviors that are larger than the sum of the parts
(Turvey, 2018). When analyzed as dynamical systems, many phenomena
exhibit new nonlinear patterns that were formerly overlooked. Continuous
data subject to linear analyses becomes flattened, and the variability
that is fundamental to these systems and their evolution over time is
ignored. Nonlinear methods provide means for capturing the rich temporal
dynamics and variability that unfold in continuous data.

The primary methods used in the current work is cross-recurrence
quantification analysis (CRQA; Zbilut, Giuliani, \& Webber Jr, 1998) and its offshoot,
windowed cross-recurrence quantification analysis (WCRQA; Webber Jr \& Zbilut, 2005).
CRQA has successfully been used to gain a better understanding of many
complex human phenomena, such as human social interactions
(for a review, see Fusaroli, Konvalinka, \& Wallot, 2014) and interpersonal synchrony
(Paxton \& Dale, 2017). We argue that the use of nonlinear methods in
the investigation of political science phenomena will lead to new insights
and development of more robust theories.

@@

The corpus of tweets used included 697428 Syrian tweets
occurring between March 30, 2012 and June 15, 2012. Using the Twitter
streaming API, servers at the University of Texas at Austin collected
the tweets. The servers were run on a stable (LTS) version of Ubuntu
Linus, 10.04, and were connected directly to the high-bandwidth UT network
maintained by IT services.

Tweets collected contained one or more of the following keywords:
\enquote{syria,} \enquote{syrian,} \enquote{damascus,} \enquote{homs,} \enquote{al-assad,} and \enquote{sunni.} These
keywords were chosen based on the most distinctive terms used in news
articles covering the Arab Spring in Syria at the time. A node.js script
was used to reinitialize the connection to Twitter's servers whenever the
script timed out due to a period of silence. The script requested data from
the API endpoint using basic HTTP authentication (OAuth was optional, at
the time) at \url{https://stream.twitter.com/1.1/statuses/filter.json}, and stored
the data directly to the local hard disk. The tweets were then post-processed
with a Python script to convert the JSON-formatted tweets to tab-separated
files and were then compressed to maximize storage efficiency.

All scripts used for scrapping tweets and compressing data can be found at \url{https://github.com/chbrown/twilight}.

@@

The social cohesion or linguistic similarity of daily tweets was calculated
by quantifying the frequency of shared content words with respect to those
words' frequencies in the entire tweet corpus. If a linguistic term has a
high frequency in one tweet but has a low frequency in the whole tweet
corpus, then it is more likely a content word and is this given more weight
in calculating social cohesion.

The tweets were first sorted into ascending order by timestamp. They were
then grouped into successive sets of five tweets. Inverse frequency
weighting for the overlap words and all words was conducted for each pair
of two tweets in the group of five, then divided to get the social cohesion
score for that pair (see Eq. 1-3). This resulted in ten social cohesion
values for each group of five tweets, which were then averaged. Finally,
the mean of all average cohesion values for all groups was taken to get
a social cohesion score for each day.

@@

Using the Integrated Crisis Early Warning System (ICEWS) developed at the
Defense Advanced Research Project Agency, the daily events in Syria were
scored with an intensity rating ranging from -10 to 10. Negative events,
such as riots and attacks, are coded with low negative numbers while
positive, peaceful events are coded with high positive numbers. This
well-validated coding scheme allows for numerical quantification of
different event types.

For both the target only and the target and source data, time series of
the count of all events were generated by summing the total number of events
on each day, irrespective of ICEWS score. Similarly, time series of count
of positive events were generated by summing the number of ICEWS events
for each day that had an intensity rating greater than zero. Lastly,
time series of count of negative events were generated by summing the number
of ICEWS events for each day that had an intensity rating less than zero.

@@

We first used cross-recurrence quantification analysis
(CRQA; Zbilut et al., 1998) to investigate the temporal patterns of
social cohesion and real-world events in Syria during Arab Spring. As
an extension of RQA, which captures the patterns of revisited states of
a single time series (Webber Jr \& Zbilut, 1994; Zbilut \& Webber Jr, 1992), CRQA
allows for two time series to be projected onto one another to investigate
shared states and trajectories. Nonlinear analyses such as CRQA can reveal
dynamics that are not captured by traditional linear methods.

In order to perform CRQA, the time series need to be converted to the same
scale, while still preserving a sampling rate that respects the native
timescale of the system
(cf.~Nyquist frequency in signal processing; Grenander, 1959).
We converted each time series (e.g., Figure \ref{fig:raw-ts}) into deciles
(e.g., Figure \ref{fig:deciled-ts}) to analyze the dynamics of different
levels of social cohesion and counts of all, positive, and negative events
on each day. In CRQA, a recurrent point is when two signals share the same
state. These shared states can be at the same time or at later points in
the time series.

@@

The percentage of recurrent points to total possible points is a metric
called recurrence rate (RR). This metrics gives a general description of
how frequently the time series are sharing the same states. In the current
study, a recurrent point would indicate that the level of social cohesion
and frequency of events were at the same intensity.

Recurrent points that occur in successive moments form line structures.
These are conceptualized as shared trajectories, where the two time series
are moving together across time in a shared state. The percentage of
recurrent points that make up diagonal line structures (i.e., two or more
consecutive points) is a metric called percent determinism (DET). For the
current project, a high DET would indicate that not only are the levels of
social cohesion and frequency of events occurring at similar intensities,
but they are doing so across consecutive days.

The average length of the line structures is called mean line (L). This can
be conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called
max line (maxL). maxL indicates the longest instance of being \enquote{stuck} in a
particular state together. The total number of lines is called NRLINE. The
more lines there are, the more coordinated the system may be. The
complexity or variability of line lengths is captured by a metric called
entropy (ENTR) and entropy while accounting for the total number of lines in
the plot is normalized entropy (rENTR). These entropy metrics help indicate
the stability and predictability of the system. When one of the signals gets
\enquote{stuck} in a state and the other moves on to other states, this results in
vertical line structures. The proportion of points forming vertical line
structures is a metric called laminarity (LAM) and the average length of the
vertical trajectories is called trapping time (TT). For more detail on CRQA
metrics, see Coco and Dale (2018).

@@

The time period of differing dynamics, from observation 38 through 43,
appears to be captured by WCRQA of social cohesion and count of negative
events for both RR and DET (Figure \ref{fig:plot-wcrqa-targ-all}). The
amount of recurrence and determinism for these periods almost consistently
lie above chance, suggesting that the intensity of negative events coincide
with the degree of social cohesion found on Twitter, and that these states
carry across multiple days in a way that is different from that of chance.

The above-chance levels of determinism and recurrence rate are demonstrating
the coupling of social cohesion and counts of all and negative events during
times of strife. We can see this string of highly violent events reflected
in the mode ICEWS event for each day (Figure \ref{fig:plot-mode-event}),
where from May 1st, 2020 to May 15th, 2020, the most frequently occurring
event was almost consistently -10, the most violent event in the coding
system. These dates directly correlate with the above chance RR and DET
metrics for the WCRQA of social cohesion and count of negative events.
Furthermore, for WCRQA of social cohesion and positive event count during
this time period, DET was entirely within chance (Figure
\ref{fig:plot-wcrqa-targ-all}). This supports our hypothesis, as the
positive events are seemingly not coupled with the social cohesion on Twitter
during civil uprisings during the Arab Spring.

@@
